Supporting Information : Effect of copula on extinction risk

### 2019-03-13 MainText.Rmd
writing methods...continued, extrisk for 25 years

<!--stochastic matrix model : age structured-->
To investigate how copula structure in noisy environment affect the extinction risk of
populations in a single patch age-structured model,
we considered a simple stochastic age-structured model (Eq. \ref{eq_age_str}) 
with two age classes: 
one is egg ($E$) and the other is adult ($A$). We assumed
the population was semelperous, i.e., adult ($A$) went through the
reproductive episode only once before death. Eggs and adults in the 
next generation depend upon the fecundity of adult ($f_a \geq 0$) 
and survival of eggs ($0 \leq s_e \leq 1$) from present generation, respectively.
The environmental stochasticity in the model was embedded in 
the fluctuating vital rates ($f_a$ and $s_e$). We considered the fecundity ($f_a$)
and survival ($s_e$) rates were random draw from a gamma ($\Gamma(k,\theta)$) and
a beta ($\mathcal{B}(\alpha,\beta)$) distribution, respectively,
which were generated from a specific tail-dependent copula
(for details see Appendix \ref{SM-age_str_vitalrates}).
Eq. \ref{eq_vitalrates} represent the stochasticity in the vital rates.
Demographic stochasticity was also taken into consideration when we included the 
density dependence effect in one of the vital rate, namely, fecundity ($f_a$).
If $\Omega = \infty$ in Eq. \ref{eq_vitalrates}, the projection matrix was independent
of populations from any age-class and the model was then called to be "linear", whereas
its "nonlinear" version had finite $\Omega$.

@@

<!--calculation of ext risk with age-structured model-->

>>>>

<!--stochastic matrix model : age structured-->
To investigate how copula structure in noisy environment affect the extinction risk of
populations in a single patch age-structured model,
we considered a simple stochastic age-structured model (Eq. \ref{eq_age_str}) 
with two age classes: 
one is egg ($E$) and the other is adult ($A$). We assumed
the population was semelperous, i.e., adult ($A$) went through the
reproductive episode only once before death. Eggs and adults in the 
next generation depend upon the fecundity of adult ($f_a \geq 0$) 
and survival of eggs ($0 \leq s_e \leq 1$) from present generation, respectively.
The environmental stochasticity in the model was embedded in 
the fluctuating vital rates ($f_a$ and $s_e$). We considered the fecundity ($f_a$)
and survival ($s_e$) rates were random draw from a gamma ($\Gamma(k,\theta)$) and
a beta ($\mathcal{B}(\alpha,\beta)$) distribution, respectively. These distribution 
could be generated from iid noises showing specific tail-dependence (left or right tail)
between fecundity and survival rates. Eq. \ref{eq_vitalrates} represent the stochasticity 
in the vital rates. Demographic stochasticity was also taken into consideration when 
we included the density dependence effect in one of the vital rate, namely, fecundity ($f_a$).
If $\Omega = \infty$ in Eq. \ref{eq_vitalrates}, the projection matrix was independent
of populations from any age-class and the model was then called to be "linear", whereas
its "nonlinear" version had finite $\Omega$.

@@

<!--calculation of ext risk with age-structured model-->
To calculate extinction risk after each time step, we simulated model (\ref{eq_age_str}) for $25$ years
with $10000$ simulations. For each simulation, we started with initial population $(E,A)_{t=0} = (5,10)$
and following Eq. \ref{eq_age_str} after each time step checked the total population $(E+A)_{t}$
was below a certain extinction threshold (we set it as $2$) or not. If it was below that threshold then we 
considered population went extinct for that simulation. We also calculated $2.5^{th}$ and $97.5^{th}$ quantiles
of $10000$ simulations with the probability of computed extinction risk at each time step.

We repeated the above analysis considering the vital rates coming from a 
an asymmetric copula (Appendix \ref{SM-age_str_vitalrates}) with specific 
left (Clayton, Survival Gumbel)
or right (Survival Clayton, Gumbel) tail dependence
or from an extreme tail-depentdent one, when vital rates were considered to be affected
by extreme climatic events.

### 2019-03-13 SuppMat.Rmd
started writing methods
writing methods...continued, extrisk for 25 years

numsims<-10000
numsteps<-100
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-Inf
ktau<-0.6  # kendall's tau

@@

```{r fig_agestr_C_SC_G_SG_linear, echo=F,results="hide"}
xL<-C_age$extdf[1:26,]
xR<-SC_age$extdf[1:26,]
t<-c(0:(nrow(xL)-1))
pdf(paste("./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf",sep=""),height=5,width=5)
op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
x<-xL
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")
x<-xR
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
lines(t,x$erA,col="blue",type="l")
legend("topleft", c("vital rates : LTdep., C","vital rates : RTdep., SC"), lty=c(1,1),  
       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
par(op)
dev.off()

xL<-SG_age$extdf[1:26,]
xR<-G_age$extdf[1:26,]
t<-c(0:(nrow(xL)-1))
pdf(paste("./Results/age_str_results/extrisk_age_str_SG_G_linear.pdf",sep=""),height=5,width=5)
op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
x<-xL
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")
x<-xR
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
lines(t,x$erA,col="blue",type="l")
legend("topleft", c("vital rates : LTdep., SG","vital rates : RTdep., G"), lty=c(1,1),  
       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
par(op)
dev.off()
```

@@

numsims<-10000
numsteps<-100
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-100
ktau<-0.6  # kendall's tau

@@

```{r fig_agestr_C_SC_G_SG_nonlinear, echo=F,results="hide"}
xL<-C_age$extdf[1:26,]
xR<-SC_age$extdf[1:26,]
t<-c(0:(nrow(xL)-1))
pdf(paste("./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_",omg,".pdf",sep=""),height=5,width=5)
op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
x<-xL
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")
x<-xR
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
lines(t,x$erA,col="blue",type="l")
legend("topleft", c("vital rates : LTdep., C","vital rates : RTdep., SC"), lty=c(1,1),  
       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
par(op)
dev.off()

@@

```{r sim_agestr_with_extremetail_linear, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("ext_risk_cop_age_str.R")
source("ExtremeTailDep.R")
numsims<-10000
numsteps<-100
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-Inf

@@

```{r fig_agestr_with_extremetail_linear, echo=F,results="hide"}
xL<-las_exL$extdf[1:26,]
xR<-las_exR$extdf[1:26,]
t<-c(0:(nrow(xL)-1))
pdf(paste("./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf",sep=""),height=5,width=5)
op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
x<-xL
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")
x<-xR
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
lines(t,x$erA,col="blue",type="l")
legend("topleft", c("vital rates : LTdep.","vital rates : RTdep."), lty=c(1,1),  
       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
par(op)
dev.off()
```

```{r sim_agestr_with_extremetail_nonlinear, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("ext_risk_cop_age_str.R")
source("ExtremeTailDep.R")
numsims<-10000
numsteps<-100
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-100
# For extreme right tail dep.
mr<-retd(n = (numsteps*numsims),d = 2,rl =1,mn=0,sdev=1)
cop<-pnorm(mr) # to make normal to uniformly distributed
 tempo<-paste("./Results/age_str_results/extreme_right_taildep_nonlinear")
if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
pdf(paste(resloc,"extreme_right_tail_cop_omg_",omg,".pdf",sep=""),height = 5,width=5)
plot(cop[,1],cop[,2])
dev.off()
nlas_exR<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(nlas_exR,paste(resloc,"sim_age_str_extreme_right_taildep.RDS",sep=""))
 # For extreme left tail dep.
ml<-retd(n = (numsteps*numsims),d = 2,rl =-1,mn=0,sdev=1)
cop<-pnorm(ml) # to make normal to uniformly distributed
 tempo<-paste("./Results/age_str_results/extreme_left_taildep_nonlinear")
if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
pdf(paste(resloc,"extreme_right_tail_cop_omg_",omg,".pdf",sep=""),height = 5,width=5)
plot(cop[,1],cop[,2])
dev.off()
nlas_exL<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(nlas_exL,paste(resloc,"sim_age_str_extreme_left_taildep.RDS",sep=""))
```

```{r fig_agestr_with_extremetail_nonlinear, echo=F,results="hide"}
xL<-nlas_exL$extdf[1:26,]
xR<-nlas_exR$extdf[1:26,]
t<-c(0:(nrow(xL)-1))
pdf(paste("./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_",omg,".pdf",sep=""),height=5,width=5)
op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
x<-xL
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")
x<-xR
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
lines(t,x$erA,col="blue",type="l")
legend("topleft", c("vital rates : LTdep.","vital rates : RTdep."), lty=c(1,1),  
       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
par(op)
dev.off()
```

@@

>>>>

numsims<-10000
numsteps<-25
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-Inf
ktau<-0.6  # kendall's tau

@@

```{r fig_agestr_C_SC_G_SG_linear, echo=F,results="hide"}
xL<-C_age$extdf
xR<-SC_age$extdf
t<-c(0:(nrow(xL)-1))
pdf(paste("./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf",sep=""),height=5,width=5)
op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
x<-xL
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")
x<-xR
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
lines(t,x$erA,col="blue",type="l")
legend("topleft", c("vital rates : LTdep., C","vital rates : RTdep., SC"), lty=c(1,1),  
       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
par(op)
dev.off()

xL<-SG_age$extdf
xR<-G_age$extdf
t<-c(0:(nrow(xL)-1))
pdf(paste("./Results/age_str_results/extrisk_age_str_SG_G_linear.pdf",sep=""),height=5,width=5)
op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
x<-xL
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")
x<-xR
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
lines(t,x$erA,col="blue",type="l")
legend("topleft", c("vital rates : LTdep., SG","vital rates : RTdep., G"), lty=c(1,1),  
       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
par(op)
dev.off()
```

@@

numsims<-10000
numsteps<-25
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-100
ktau<-0.6  # kendall's tau

@@

```{r fig_agestr_C_SC_G_SG_nonlinear, echo=F,results="hide"}
xL<-C_age$extdf
xR<-SC_age$extdf
t<-c(0:(nrow(xL)-1))
pdf(paste("./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_",omg,".pdf",sep=""),height=5,width=5)
op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
x<-xL
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")
x<-xR
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
lines(t,x$erA,col="blue",type="l")
legend("topleft", c("vital rates : LTdep., C","vital rates : RTdep., SC"), lty=c(1,1),  
       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
par(op)
dev.off()

@@

```{r sim_agestr_with_extremetail_linear, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("ext_risk_cop_age_str.R")
source("ExtremeTailDep.R")
numsims<-10000
numsteps<-25
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-Inf

@@

```{r fig_agestr_with_extremetail_linear, echo=F,results="hide"}
xL<-las_exL$extdf
xR<-las_exR$extdf
t<-c(0:(nrow(xL)-1))
pdf(paste("./Results/age_str_results/extrisk_age_str_extremetail_linear.pdf",sep=""),height=5,width=5)
op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
x<-xL
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")
x<-xR
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
lines(t,x$erA,col="blue",type="l")
legend("topleft", c("vital rates : LTdep.","vital rates : RTdep."), lty=c(1,1),  
       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
par(op)
dev.off()
```

```{r sim_agestr_with_extremetail_nonlinear, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("ext_risk_cop_age_str.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)
source("ext_risk_cop_age_str.R")
source("ExtremeTailDep.R")
numsims<-10000
numsteps<-25
p0<-c(5,10)
ext_thrs<-2
gshape<-2
gscale<-1
bshape1<-3
bshape2<-2
par_dist<-c(gshape,gscale,bshape1,bshape2)
omg<-100
# For extreme right tail dep.
mr<-retd(n = (numsteps*numsims),d = 2,rl =1,mn=0,sdev=1)
cop<-pnorm(mr) # to make normal to uniformly distributed
 tempo<-paste("./Results/age_str_results/extreme_right_taildep_nonlinear")
if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
pdf(paste(resloc,"extreme_right_tail_cop_omg_",omg,".pdf",sep=""),height = 5,width=5)
plot(cop[,1],cop[,2])
dev.off()
nlas_exR<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(nlas_exR,paste(resloc,"sim_age_str_extreme_right_taildep.RDS",sep=""))
 # For extreme left tail dep.
ml<-retd(n = (numsteps*numsims),d = 2,rl =-1,mn=0,sdev=1)
cop<-pnorm(ml) # to make normal to uniformly distributed
 tempo<-paste("./Results/age_str_results/extreme_left_taildep_nonlinear")
if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
pdf(paste(resloc,"extreme_right_tail_cop_omg_",omg,".pdf",sep=""),height = 5,width=5)
plot(cop[,1],cop[,2])
dev.off()
nlas_exL<-sim_age_str(p0=p0,ext_thrs = ext_thrs,cop=cop,par_dist = par_dist,omg=omg,
            numsims = numsims,numsteps = numsteps,ploton=T,resloc=resloc)
saveRDS(nlas_exL,paste(resloc,"sim_age_str_extreme_left_taildep.RDS",sep=""))
```

```{r fig_agestr_with_extremetail_nonlinear, echo=F,results="hide"}
xL<-nlas_exL$extdf
xR<-nlas_exR$extdf
t<-c(0:(nrow(xL)-1))
pdf(paste("./Results/age_str_results/extrisk_age_str_extremetail_nonlinear_omg_",omg,".pdf",sep=""),height=5,width=5)
op<-par(mar=c(5,4,1.5,1.5),mgp=c(2,0.7,0))
plot(0,-1,ylim=c(0,1),xlim=range(t),ylab="Extinction risk",xlab="time",cex.lab=2,cex.axis=1.5)
x<-xL
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(1,0,0,0.3), border = NA)
lines(t,x$erA,col="red",type="l")
x<-xR
polygon(c(rev(t), t), c(rev(x$CI0.975), x$CI0.025), col = rgb(0,0,1,0.3), border = NA)
lines(t,x$erA,col="blue",type="l")
legend("topleft", c("vital rates : LTdep.","vital rates : RTdep."), lty=c(1,1),  
       col=c('red', 'blue'), horiz=F, bty='n', cex=1.5)
par(op)
dev.off()
```

@@

# Vital rates in a stochastic age-structured model having specific tail-dependence \label{age_str_vitalrates}

A $N \times 2$ matrix, M, was generated for each of the left (Clayton, Survival Gumbel) and right 
(Survival Clayton, Gumbel) tail-dependent bivariate copulas, where $N = numsteps(=25) \times numsims(=10000)$.
We kept the same Kendall's correlation ($\tau = 0.6$) for each bivariate copula which used parameter accordingly
(see `iTau` function from `copula` package). For fecundity of adults, we transformed the first column of matrix 
$M$ as gamma ($\Gamma(k,\theta)$) distributed, $k = 2$ and $\theta = 1$ were shape and scale parameters, 
respectively. For a given $\theta$, increasing 
$k$ indicates distribution will shift towards higher fecundity which might 
reduce the extinction risk if the other parameters
remain unchanged. For survival rates of eggs, we took the second column of $M$ and applied a 
beta ($\mathcal{B}(\alpha,\beta)$) transformation, $\alpha =3$ and
$\beta = 2$ were the shape parameters which could control the distribution 
of the survival rates of eggs. Entries within each clumn of matrix $M$, were thus independent of each others but 
the two columns of $M$ had a specified Kendall's correlation ($\tau = 0.6$) and underlying tail-dependence. 
For each $25$ timesteps from each of $10000$ simulations, the vital rates (fecundity and survival rates) 
of the stochastic projrction matrix in model (\ref{MT-eq_age_str}) were drawn pairwise from each row of matrix $M$.

For extreme tail-dependence algorithm,
